---
title: "Examples for package barplotOR"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{main}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r main00}
library("barplotOr")
set.seed(1234)
```

## Simple example

```{r main01, fig.width=7}
dataset <- matrix(sample(1:20, 13*8, replace = TRUE), ncol = 13)
myCol <- c("#1B9E77", "#D95F02", "#7570B3", "#E7298A",
           "#66A61E", "#E6AB02", "#A6761D", "#666666")
barplot  (height = dataset, col = myCol, names.arg = LETTERS[1:13])
barplotOr(height = dataset, col = myCol, names.arg = LETTERS[1:13])
```

## Adding value or rank in the barplot

```{r main02, fig.width=7, fig.height=5}
dataset <- matrix(sample(10:20, 13*8, replace = TRUE), ncol = 13)
myCol <- c("#1B9E77", "#D95F02", "#7570B3", "#E7298A",
           "#66A61E", "#E6AB02", "#A6761D", "#666666")
# number in the barplot
aBarPlot <- barplotOr(
  height = dataset,
  col = myCol, decr = FALSE,
  names.arg = LETTERS[1:13])
trash <- lapply(1:ncol(dataset), function(j){
  sapply(1:nrow(dataset), function(i){
    text(
      x = aBarPlot[j],
      y = cumsum(dataset[order(dataset[,j]), j])[i],
      labels = dataset[order(dataset[,j])[i], j],
      pos = 1
    )
  })
})
# position in the barplot
aBarPlot <- barplotOr(
  height = dataset,
  col = myCol, decr = FALSE,
  names.arg = LETTERS[1:13])
trash <- lapply(1:ncol(dataset), function(j){
  sapply(1:nrow(dataset), function(i){
    text(
      x = aBarPlot[j],
      y = cumsum(dataset[order(dataset[,j]), j])[i],
      labels = rev(rank(
        dataset[order(dataset[,j]), j]))[i],
      pos = 1
    )
  })
})

# number in the barplot
aBarPlot <- barplotOr(
  height = dataset,
  col = myCol, decr = TRUE,
  names.arg = LETTERS[1:13])
trash <- lapply(1:ncol(dataset), function(j){
  sapply(1:nrow(dataset), function(i){
    text(
      x = aBarPlot[j],
      y = cumsum(dataset[order(dataset[,j], decreasing = TRUE), j])[i],
      labels = dataset[order(dataset[,j], decreasing = TRUE)[i], j],
      pos = 1
    )
  })
})
# position in the barplot
aBarPlot <- barplotOr(
  height = dataset,
  col = myCol, decr = TRUE,
  names.arg = LETTERS[1:13])
trash <- lapply(1:ncol(dataset), function(j){
  sapply(1:nrow(dataset), function(i){
    text(
      x = aBarPlot[j],
      y = cumsum(dataset[order(dataset[,j], decreasing = TRUE), j])[i],
      labels = rev(rank(
        dataset[order(dataset[,j], decreasing = TRUE), j]))[i],
      pos = 1
    )
  })
})
```

## Example with real data: counting letter occurrences

In this example occurrences of letters are extracted from the main page of Wikipedia in different languages. Please note that this is just an example to illustrate the use of barplotOr, do not interprate the results as some letters have been modified to reduce the complexity of this vignette. They do not represent the real occurrences of letters.

```{r main03, fig.width=7, fig.height=5}
# occurrences of letters
getOccuLet <- function(myURL){
  xx <- try(readLines(con = myURL))
  xx <- paste(xx, collapse = "")
  xx <- tolower(xx)
  xx <- trimws(xx)
  xx <- gsub(pattern = "à|â|ã|ä|å", replacement = "a", x = xx)
  xx <- gsub(pattern = "é|ê|è", replacement = "e", x = xx)
  xx <- gsub(pattern = "ð|ø|ù|û|ò|ò|ÿ|ó|á|ü|ú|ì|þ|î|ï|ë|ô|õ|ö|æ|í|ß|ç", replacement = "", x = xx)
  xx <- gsub(pattern = "ñ", replacement = "n", x = xx)
  xx <- gsub(pattern = "[^[:alpha:]]", replacement = "", x = xx)
  xx <- strsplit(xx, "")[[1]]
  return(table(xx))
}
getFR <- as.data.frame(getOccuLet(myURL = "https://fr.wikipedia.org/"))
getES <- as.data.frame(getOccuLet(myURL = "https://es.wikipedia.org/"))
getIT <- as.data.frame(getOccuLet(myURL = "https://it.wikipedia.org/"))
getEN <- as.data.frame(getOccuLet(myURL = "https://en.wikipedia.org/"))
getPT <- as.data.frame(getOccuLet(myURL = "https://pt.wikipedia.org/"))
getDE <- as.data.frame(getOccuLet(myURL = "https://de.wikipedia.org/"))
getPL <- as.data.frame(getOccuLet(myURL = "https://pl.wikipedia.org/"))

all <- list(getFR, getES, getIT, getEN, getPT, getDE, getPL)
res <- Reduce(function(x, y) merge(x, y, by = "xx"), all)
myCol <- sample(x = colors(), size = 26)

# with barplot:
barplot(
  height = as.matrix(res[,2:ncol(res)]),
  col = myCol, width = 1,
  names.arg = c("FR", "ES", "IT", "EN", "PT", "DE", "PL"))
legend("topright", legend = res[,1], fill = myCol, ncol = 6)

# with barplotOr:
barplotOr(
  height = as.matrix(res[,2:ncol(res)]),
  col = myCol, width = 1,
  names.arg = c("FR", "ES", "IT", "EN", "PT", "DE", "PL"))
legend("topright", legend = res[,1], fill = myCol, ncol = 6)
```

## Example from the literature

An earlier version of the barplotOr function was used to build Figure 3 in Rebaudo and Rabhi (Rebaudo F., Rabhi V.B. (2018) Modeling temperature-dependent development rate and phenology in insects: review of major developments, challenges, and future directions. Entomologia Experimentalis et Applicata, 166(8): 607-617. [DOI: 10.1111/eea.12693]).
